{\helveticaitalic{Supplementary Material}}

### 2022-02-04 tex/sections/results.tex
Split latex document into subfiles

\onecolumn
\firstpage{1}

\title[Scalar Poincaré map for bursting]{A scalar Poincaré map for anti-phase bursting in coupled inhibitory neurons with synaptic depression.}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}
\section{}
Short-term synaptic plasticity is found in many areas of the central
nervous system.  In the inhibitory half-centre central pattern
generators involved in locomotion, synaptic depression is believed to
act as a burst termination mechanism, allowing networks to generate
anti-phase bursting patterns of varying periods.  To better understand
burst generation in these central patter generators, we study a
minimal network of two neurons coupled through depressing synapses.
Depending on the strength of the synaptic conductance between the two
neurons, this network can produce symmetric $n-n$ anti-phase bursts,
where neurons fire $n$ spikes in alternation, with the period of such
solutions increasing with the strength of the synaptic conductance.
Relying on the timescale disparity in the model, we reduce the
eight-dimensional network equations to a fully-explicit scalar
Poincaré burst map.  This map tracks the state of synaptic depression
from one burst to the next and captures the complex bursting dynamics
of the network.  Fixed points of this map are associated with stable
burst solutions of the full network model, and are created through
fold bifurcations of maps.  We derive conditions that describe
period-increment bifurcations between stable $n-n$ and $(n+1)-(n+1)$
bursts, producing a full bifurcation diagram of the burst cycle
period.  Predictions of the Poincaré map fit excellently with
numerical simulations of the full network model and allow the study of
parameter sensitivity for rhythm generation.

\tiny
 \keyFont{ \section{Keywords:} Synaptic depression, Poincaré map, Dynamical system, Neuronal bursting, Central pattern generator}

\end{abstract}

\section{Introduction}
% DEPRESSION
Short-term synaptic plasticity may have a role in burst activity in
central pattern generators (CPGs).  Short-term synaptic depression is
commonly found in neuronal networks involved in the generation of
rhythmic movements, such as in the pyloric CPG of the spiny
lobster~\citep{manor1997, rabbah2007}, or in the lumbosacral cord of
the chick embryo~\citep{donovan1998}.  Synaptic depression modulates
the strength of synapses in response to changes to the presynaptic
firing frequency.  At a high neuronal firing frequency, depression
weakens the strength of synapses and therefore reduces the magnitude
of the postsynaptic response.  At low firing frequency, it allows
sufficient time for the synapse to recover from depression between
spikes, leading to a stronger postsynaptic response.  In reciprocal
networks, synaptic depression has been shown to act as a ``switch'',
giving rise to a wide range of network dynamics such as synchronous
and multi-stable rhythms, as well as fine tuning the frequency of
network oscillations~\citep{nadim2000, nadim1999, bose2011}.

% Brown
~\citet{brown1911} pioneered the idea that synaptic depression acts as
a burst termination mechanism in reciprocally inhibitory CPGs involved
in rhythm generation of locomotion. When one side is firing during a
burst the other, antagonistic side, is prevented from firing by
synaptic inhibition. However, the weakening of inhibition as a result
of synaptic depression eventually releases the antagonistic side so
that it starts firing, terminating the burst on the side that had
originally been firing.  This rhythmogenesis hypothesis has been
considered one of a handful of standard mechanisms for generating
locomotion rhythms in
vertebrates~\citep{reiss1962,perkel1974,friesen1994}. It has been
proposed as an explaination of the antiphase burst rhythm in
struggling in \textit{Xenopus} tadpoles~\citep{li2007}.

~\citet{bose2011} investigated burst generation in a generic
half-centre CPG that consists of two identical, tonically active
Morris-Lecar~\citep{morris1981} neurons coupled through inhibitory
depressing synapses.  Numerical simulations showed that when the
reciprocal synaptic conductance between the two neurons is varied, the
network produces symmetric $n-n$ anti-phase bursts, with stronger
synaptic coupling leading to longer bursts.  They used methods from
geometric singular perturbation theory to separate the timescales of
the fast membrane, and the slow synaptic dynamics of the network to
derive one-dimensional conditions necessary for the existence of
stable $n-n$ solutions (for $n\leq 2$). According to these conditions
the type of firing pattern largely depends on the slow depression
dynamics of the synapses between the two neurons, and can therefore be
predicted by knowing the strengths of the synaptic conductances of the
two synapses.
Thus, the scalar conditions derived in~\citet{bose2011} provide a
method to numerically identify the type of stable $n-n$ pattern for
any given value of the coupling strength.  However, they do not
predict the exact period of such solutions.
Furthermore, while they provide good arguments for the validity of
their reduction assumptions and the resulting scalar conditions, they
do not verify them numerically.

Here we extend the previous analysis by providing a Poincaré map of
the slow depression dynamics.  This allows us not only to predict the
types of stable $n-n$ solutions the full network can produce, (for any
$n$), but also to study how varying the coupling strength affects the
period of such solutions.  To do this, we build on, and numerically
test, the assumptions on the fast-slow timescale disparity made
in~\cite{bose2011}.  We reduce the two-cell model to a scalar
Poincaré map that tracks the evolution of the depression from the
beginning of one burst to the beginning of the next burst.  Stable
fixed points of our map are associated with stable $n-n$ burst
solutions.  Our map construction is motivated by the burst length map
of a T-type calcium current, utilised by~\citet{matveev2007}, which
approximates the anti-phase bursting dynamics of a network of two
coupled Morris-Lecar neurons.  In contrast to our model, the network
described in the~\cite{matveev2007} paper does not contain short-term synaptic
depression, and burst termination is instead accomplished through the
dynamics of a slow T-type calcium current.

% MAIN RESULTS AND CONCLUSIONS
The Poincaré map derived here replicates the results from numerical
simulations of the full two-cell ODE system: Given the strength of
maximum conductance between the two neurons, fixed points of our map
predict the type and period of $n-n$ patterns, the switch between
burst solutions of different periods, as well as the occurrence of
co-existent solutions.  In addition to proving the existence and
stability of fixed points, our map shows that fixed points are created
via a fold bifurcation of maps.  We also derive algebraic conditions
that predict the period increment bifurcations that allow the switch
between $n-n$ and $(n+1)-(n+1)$ solutions.  Because our map is fully
explicit, it lays the framework for studying the effects of other
model parameter on network dynamics without the need to run expensive
numerical integrations of the ODEs.

% PAPER STRUCTURE.
This paper is organised as follows.  First, we introduce the network
of two neurons, and describe the properties of single cell and synapse
dynamics.  We use numerical simulations of the network to provide an
intuition for the range of possible burst dynamics the system can
produce.  Next, we state and justify the simplifying assumptions that
are necessary for the map construction.  Finally, we analytically
derive the first return map of the depression variable as well as the
conditions that are required for stable solutions to bifurcate between
$n-n$ and $(n+1)-(n+1)$ solutions.  We end this work with a
discussion.

\section{Materials and Methods}
We consider a pair of identical Morris-Lecar neurons~\citep{morris1981}, with parameters adapted from~\cite{bose2011}.
The Morris-Lecar model is a set of two first-order differential equations that describe the membrane dynamics of a spiking neuron.
The depolarisation is modelled by an instantaneous calcium current, and the hyperpolarisation by a slow potassium current and a leak current.
The membrane potential $v_{i}$ and potassium activation $w_{i}$ of neuron $i$ ($i, j=1,2$) is described by:
\begin{align}
    ~\label{eq:cell-modelA}
     \dot v_{i}&= f(v_{i}, w_{i}) -\gbar s_j(v_i-v_{s}),\\
    ~\label{eq:cell-modelB}
     \dot w_{i} &=h(v_i,w_i).
   \end{align}
Here $v_{s}$ is the inhibitory reversal potential, and $\gbar$ and $s_{j}$ are the maximal synaptic conductance and the synaptic gating, respectively, constituting the total inhibitory conductance $\gbar s_{j}$ from neuron $j$ to neuron $i$.
Function $f(v_{i}, w_{i})$ describes the membrane currents of a single cell:
\begin{equation}
 ~\label{eq:memcur}
  f(v_{i}, w_{i}) = -g_{\ca}m_{\infty}(v_{i})(v_{i}-v_{\ca}) - g_{\k}w_{i}(v_{i}-v_{\k})-g_{\leak}(v_{i}-v_{\leak}) + I.
\end{equation}
The currents include a constant current $I$, and three ionic currents:
an instantaneous calcium current, a potassium current, and a leak current, with respective reversal potentials $v_{\ca}$, $v_{\k}$, and $v_{\leak}$, as well as maximum conductances $g_{\ca}$, $g_{\k}$, and $g_{\leak}$.
The function $h(v_{i}, w_{i})$ models the kinetics of the potassium gating variable $w_{i}$, and is given by
\begin{equation}
  h(v_{i}, w_{i})=\frac{w_{\infty}(v_{i})-w_{i}}{\tau_{w}}~\label{eq:h}.
\end{equation}
The steady-state activation functions $m_{\infty}$ and $w_{\infty}$ as well as the default model parameters are described in the \textbf{Supplementary Material S1}.

The dynamics of the synaptic interactions between the neurons are governed by a synaptic gating variable $s_{i}$ and a depression variable $d_{i}$:
\begin{linenomath}
  \begin{align}
    \dot s_{i} &= -s_{i}/\tau_{s}~\label{eq:dot-s},\\
    \dot d_{i} &= (1-d_{i})/\tau_{d}~\label{eq:dot-d}.
  \end{align}
\end{linenomath}
Variable $d_{i}$ describes a firing rate dependent depletion mechanism that governs the amount of depression acting on the synapse.
Our model is agnostic with respect to the exact mechanism of this depletion, be it pre- or post-synaptic.
In between spikes $s_{i}$ decays with time constant $\tau_{s}$, while $d_{i}$ recovers with time constant $\tau_{d}$.
Because synaptic depression occurs on a much slower timescale than synaptic inhibition, we assume $\tau_{d}\gg\tau_{s}$.
At spike time when the voltage $v_{i}$ crosses a synaptic threshold $v_{\theta}$, the synaptic variable $s_{i}$ is reset to the current value of $d_{i}$, and $d_{i}$ is scaled down by the depression strength parameter $\lambda$, where $0<\lambda<1$.
This results in the following discontinuous reset rule on spike time:
\begin{linenomath}
  \begin{align}
    s_{i} &\leftarrow d_{i}~\label{eq:s-reset},\\
    d_{i} &\leftarrow \lambda d_{i}~\label{eq:d-reset}.
  \end{align}
\end{linenomath}
The equations for the depression model were adapted from the~\citet{bose2001} model.
These equations are a mathematically tractable simplification of the established phenomenological depression model previously described by~\citet{tsodyks1997}.
The original equations in~\cite{bose2001} include an exponential decay of $d_{i}$ throughout the active phase of the action potential, that is $\dot d_{i}=-d_{i}/\tau_{\beta}$ for $v_{i}>v_{\theta}$.
To simplify the mathematical analysis we have reduced these dynamics to a single-reset event at spike time according to \cref{eq:d-reset}.
This simplification is natural given an approximately constant duration of the active phase of an action potential, and it will become evident from our results that this simplification does not qualitatively change the dynamics of the network compared to the original equations, where depression dynamics are described by an exponential decay.

When the cells are uncoupled ($\gbar=0$), the membrane dynamics are determined by the cubic $v$-nullcline $v_{\infty}(v_i)$ and the sigmoid $w$-nullcline $w_{\infty}(v_{i})$, satisfying $\dot v_{i}=0$ and $\dot w_{i}=0$, respectively.
The two curves intersect along the middle branch of $v_{\infty}$, creating an unstable fixed point $p_{f}=(v_{f},w_{f})$ with a surrounding stable limit cycle of period $T$ (\cref{fig:nullclines}A).
Trajectories along that limit cycle have the familiar shape of the action potential (\cref{fig:nullclines}B).
The trajectory of an action potential can be dissected into four phases: (1) a silent phase, (2) a jump up, (3) an active phase, and (4) a jump down~\citep[see e.g.][]{ermentrout2010}.
During the silent phase the trajectory evolves along the left branch ($v_{i}<v_{\theta}$) of the cubic $v$-nullcline.
Once the trajectory reaches the local minimum of $v_{\infty}$, it ``jumps up'' to the right branch ($v_{i}>v_{\theta}$), crossing the firing threshold $v_{\theta}$.
During the active phase the trajectory then evolves along the right branch of the cubic until it arrives at the local maximum, where it ``jumps down'' to the left branch commencing a new cycle.

The two-cell network model is numerically integrated using an adaptive step-size integrator for stiff differential equations implemented with XPPAUT~\citep{ermentrout2002} and controlled through the Python packages SciPy~\citep{scipy2020} and PyXPP~\citep{pyxpp}.
The following mathematical analysis is performed on the equations of a single cell.
Unless required for clarity, we will therefore omit the subscripts $i,j$ from here on.

\section{Results}
\subsection{Anti-phase burst solutions}
Short-term synaptic depression of inhibition in a half-centre oscillator acts as a \emph{burst termination} mechanism~\citep{brown1911} and is known to produce $n-n$ anti-phase burst solutions of varying period.
Such $n-n$ solutions consist of cells firing bursts of $n$ spikes in alternation.
\Cref{fig:depression-traces}A shows the timecourse of a typical $4-4$ burst.
While one cell is firing a burst it provides an inhibitory conductance to the other cell,  preventing it from firing.
Therefore, at any given moment one cell is spiking while the other is inhibited.
Consistent with~\citet{bose2011} we will refer to the currently firing cell as ``free'' and we will call the inhibited cell ``quiet''.
Additionally, we will distinguish between two phases of a $n-n$ solution:
We will refer to the burst duration of a cell as the ``free phase'', which is the time between the first spike and the last spike in a burst. And we will call the remaining duration of a cycle, when a cell is not spiking, the ``quiet phase''.

@@

Suppose that at $t=0$ cell 1 becomes free with some initial $d(0)$. 
Cell 1 then fires $n$ spikes at the uncoupled period $T$.
Let $s(t)$ and $d(t)$ be the corresponding solutions of the synaptic and depression variables of cell 1.
After $n$ spikes the total conductance $\gbar s(t)$ acting on the quiet cell 2 has decayed sufficiently to satisfy the release condition~\eqref{eq:release}, that is at some time $t=(n-1)T+\delt$ we have $\gbar s(t)=g^{\star}$.
Cell 2 is then released and prevents cell 1 from further spiking.
Here $\delt$ is the \emph{inter-burst-interval}, or the time between the last spike of cell 1 and the first spike of cell 2~\citep{bose2011}.
Once released, cell 2 also fires $n$ spikes until cell 1 becomes free once again at the cycle period.
Let $P_n$ denote the full cycle period of a $n-n$ solution:
\begin{equation}
 ~\label{eq:P}
  P_n = 2(n-1)T + 2\delt.
\end{equation}
We can now define the free and quiet phases of cell 1 explicitly. The free phase is the time interval between the first and last spikes of the burst, that is for time $0<t<(n-1)T$.
During the free phase of cell 1, the quiet cell 2 is inhibited sufficiently strong to prevent it from firing, hence $\gbar s > g^{\star}$.
The quiet phase of cell 1 is the remaining duration of the cycle when the cell is not firing, that is for $(n-1)T < t < 2(n-1)T + 2\delt$.

Note that only the quiet phase depends on $\delt$, and $\delt$ will play a central role in the construction of $\Pi_{n}$.
From \cref{eq:P} $\delt$ can be be computed as
\begin{equation}
 ~\label{eq:delta-t-P}
  \delt = \frac{1}{2}P_n - (n-1)T.
\end{equation}
\noindent
We can use \cref{eq:delta-t-P} and the numerically computed bifurcation diagram of the period for stable $n-n$ solutions in \cref{fig:bif-diagram}A to obtain the graph of $\delt$ as a function of $\gbar$ (\cref{fig:delta-t}).
Each continuous branch of $\delt$ is monotonically increasing and corresponds to a $n-n$ burst:
Stronger coupling $\gbar$ increases the total synaptic conductance $\gbar s$ that acts on the quiet cell, thus delaying its release.
It is easy to see that for any $n$-branch we have $\delt<T$:
Once $\delt$ crosses $T$, the free cell can ``squeeze in" an additional spike and the solutions bifurcate into a $(n+1)-(n+1)$ burst.

Distinguishing between the free and quiet phases of a cycle allows us to describe the dynamics of the depression variable $d$ explicitly for each phase.
As can be seen from \cref{fig:free-quiet1}C, during the free phase $d$ depresses at spike times and recovers in the $ISI$s.
In contrast, during the quiet phase $d$ only recovers and does not depress.
Given the initial $d^{\star}=d(0)$ at the beginning of the cycle and the number of spikes in the free phase $n$, we can now construct the burst map $\Pi_{n}$.
The map
\begin{equation}
  \Pi_{n}(d^{\star})=Q_{n}\big(F_{n}(d^{\star}\big))
\end{equation}
\noindent
is a composition of two maps.
Map
\begin{equation}
  F_{n}:d^{\star}\mapsto \delt
\end{equation}
models the evolution of $d$ in the free phase.
$F_{n}$ takes an initial value $d^{\star}$ and calculates the inter-burst-interval $\delt$.
Map
\begin{equation}
  Q_{n}:\delt \mapsto d^{\star}
\end{equation}
models the recovery of $d$ in the quiet phase.
Given some $\delt$ map $Q_n$ computes $\dstar$ at the start of the next cycle.

@@

We start the construction of $\Pi_n$ by first considering the free phase and building the map $F_n$.
At each spike time $t_{k}$, variable $d$ depresses with the reset $d(t_{k})\leftarrow \lambda d(t_{k}^{-})$ (\cref{eq:d-reset}).
For the duration of the subsequent $ISI$, the recovery of $d$ is described by the solution to \cref{eq:dot-d}:
\begin{equation}
  d(t) = 1 - (1-\lambda d_{k})e^{-t/\tau_{d}}~\label{eq:d-sol},
\end{equation}
where the variable $d_{k}=d(t^{-}_{k})$ is the level of depression at spike time $t_{k}$.
By substituting $t=T$ we can build a linear map that models the depression of $d$ from spike time $t_{k}$ to the subsequent spike time $t_{k+1}$ during the free phase:
\begin{equation}
  d_{k+1} = \lambda\rho d_{k} + (1-\rho),~\label{eq:map-d}
\end{equation}
where to keep the notation simple we let $\rho=\exp(-T/\tau_{d})$.
Since $0<\lambda, \rho<1$, the map~\eqref{eq:map-d} is increasing and contracting, with a fixed point at
\begin{equation}
 ~\label{eq:dsup}
  d_{s}=\frac{1-\rho}{1-\lambda\rho},
\end{equation}
where $0<d_{s}<1$.
The value $d_{s}$ is the maximum depression value that can be observed in the suppressed solution where the active cell fires at its uncoupled period $T$ (see \cref{fig:burst-sols}E).
Using the release condition in \cref{eq:release} allows us to derive the value of the the minimum coupling strength that will produce the full suppressed solution, denoted as $\gbar_{s}$.
Solving \cref{eq:dot-s} for $s(t)$ with $t=T$ and setting the initial value $s(0)=d_{s}$ gives
\begin{equation}
\gbar_{s} d_{s}e^{-T/\tau_{s}}=g^{\star}.
\end{equation}
By further substituting the definition of $d_{s}$ in~\eqref{eq:dsup} and rearranging, we can write $\gbar_{s}$ as a function of the depression strength $\lambda$:
\begin{equation}
  \label{eq:gs}
  \gbar_{s}(\lambda) = g^{\star}e^{T/\tau_{s}}\frac{1-\lambda \rho}{1-\rho}.
\end{equation}
Note that the above dependence of $\gbar_{s}$ on $\lambda$ is linear and monotonically decreasing.
Increasing $\lambda$ reduces the strength of the depression of the free cell.
This in turn allows the free cell to fully suppress the quiet cell at smaller values of $\gbar$.

Solving~\eqref{eq:map-d} gives us the linear map $\delta_{n}:d^{\star}\mapsto d_{n}$, that for some initial $d^{\star}$ computes the depression at the $n$th spike time, $d_{n}=d(t_{n}^{-})$:
\begin{equation}
 ~\label{eq:delta-map}
  \delta_{n}(d^{\star}) = (\lambda \rho)^{n-1} d^{\star} + (1-\rho)\sum_{i=0}^{n-2}(\lambda \rho)^{i}.
\end{equation}
Since $\lambda < 1$, function $\delta_{n}$ is a linearly increasing function of $d^{\star}$ with a fixed point at $d_{s}$ for all $n$.
Having identified $d$ after $n$ spikes, we can now use the release condition $\gbar s = g^{\star}$ (\cref{eq:release}) to find $\delt$.
At the last spike of the free phase at time $t_{n}=(n-1)T$ the synapse variable $s$ is reset according to \cref{eq:s-reset}:
\begin{equation}
  s(t_{n})\leftarrow \dn(\dstar).
\end{equation}
After the last spike $s$ decays exponentially according to the solution of \cref{eq:dot-s}, where setting the initial condition $s(0)=\dn(\dstar)$ yields:
\begin{equation}
~\label{eq:s-sol}
  s(\delt)=\dn(\dstar)e^{-\Delta t/\tau_{s}}.
\end{equation}
Substituting $s(\delt)$ into $s$ of the release condition (\cref{eq:release}) gives then
\begin{equation}
 ~\label{eq:release2}
  \gbar \dn(\dstar) e^{-\delt/\tau_{s}}=g^{\star}.
\end{equation}
Our assumption of the release condition guarantees that the quiet cell 2 spikes and becomes free when $\gbar s - g^{\star}$ crosses zero.
Solving \cref{eq:release2} for $\delt$ allows us to compute the inter-spike-interval as a function of $\dstar$, which defines our map $F_n$:
\begin{equation}
 ~\label{eq:Fn-map}
  F_{n}(d^{\star}):=\tau_{s}\ln{\left(\frac{\gbar }{g^{\star}} \delta_{n}(d^{\star})\right)}= \delt.
\end{equation}

@@

From \cref{eq:Pn-map} it is evident that $\Pi_{n}$ is monotonically increasing with respect to $\gbar$ and also $\dstar$:
\begin{align}
  \label{eq:non-degen1}
  \dv{\Pi_{n}}{\gbar}&>0,\\[0.5ex]
  \label{eq:dPidd}
  \dv{\Pi_{n}}{\dstar}&>0,
\end{align}
and in the following sections we will heavily rely on this monotonicity property of $\Pi_n$.
Just as $F_{n}$, curves $\Pi_{n}$ spawn at the asymptote $d_{a}$ (\cref{eq:da}), and because
\begin{equation}
  \lim_{\gbar \to \infty}\Pi_{n} = 1\text{ for all }n,
\end{equation}
fixed points of $\Pi_{n}$ lie in $(d_{a}, 1)$.

>>>>

\subsection{Anti-phase burst solutions}
Short-term synaptic depression of inhibition in a half-centre oscillator acts as a \emph{burst termination} mechanism~\citep{brown1911} and is known to produce $n-n$ anti-phase burst solutions of varying period.
Such $n-n$ solutions consist of cells firing bursts of $n$ spikes in alternation.
\Cref{fig:depression-traces}A shows the timecourse of a typical $4-4$ burst.
While one cell is firing a burst it provides an inhibitory conductance to the other cell,  preventing it from firing.
Therefore, at any given moment one cell is spiking while the other is inhibited.
Consistent with~\citet{bose2011} we will refer to the currently firing cell as ``free'' and we will call the inhibited cell ``quiet''.
Additionally, we will distinguish between two phases of a $n-n$ solution:
We will refer to the burst duration of a cell as the ``free phase'', which is the time between the first spike and the last spike in a burst. And we will call the remaining duration of a cycle, when a cell is not spiking, the ``quiet phase''.

@@

Suppose that at $t=0$ cell 1 becomes free with some initial $d(0)$.
Cell 1 then fires $n$ spikes at the uncoupled period $T=T_{act}+T_{inact}$.
Let $s(t)$ and $d(t)$ be the corresponding solutions of the synaptic and depression variables of cell 1.
After $n$ spikes the total conductance $\gbar s(t)$ acting on the quiet cell 2 has decayed sufficiently to satisfy the release condition~\eqref{eq:release}, that is at some time $t=(n-1)T + \delt$, where $\Delta t < T_{inact}$, we have $\gbar s(t)=g^{\star}$.
Cell 2 is then released and prevents cell 1 from further spiking.
Here $\delt$ is the time between the last spike of cell 1 and the first spike of cell 2 ~\citep{bose2011}.
Once released, cell 2 also fires $n$ spikes until cell 1 becomes free once again at the cycle period.
Let $P_n$ denote the full cycle period of a $n-n$ solution:
\begin{equation}
  ~\label{eq:P}
  P_n = 2(n-1)T + 2\delt.
\end{equation}
We can now define the free and quiet phases of cell 1 explicitly. The free phase is the time interval between the first and last spikes of the burst, that is for time $0<t<(n-1)T$.
During the free phase of cell 1, the quiet cell 2 is inhibited sufficiently strong to prevent it from firing, hence $\gbar s > g^{\star}$.
The quiet phase of cell 1 is the remaining duration of the cycle when the cell is not firing, that is for $(n-1)T < t < 2(n-1)T + 2\delt$.

Note that only the quiet phase depends on $\delt$ which will play a central role in the construction of $\Pi_{n}$.
From \cref{eq:P} $\delt$ can be be computed as
\begin{equation}
  ~\label{eq:delta-t-P}
  \delt = \frac{1}{2}P_n - (n-1)T.
\end{equation}
\noindent
We can use \cref{eq:delta-t-P} and the numerically computed bifurcation diagram of the period for stable $n-n$ solutions in \cref{fig:bif-diagram}A to obtain the graph of $\delt$ as a function of $\gbar$ (\cref{fig:delta-t}).
Each continuous branch of $\delt$ is monotonically increasing and corresponds to a $n-n$ burst:
Stronger coupling $\gbar$ increases the total synaptic conductance $\gbar s$ that acts on the quiet cell, thus delaying its release.
It is easy to see that for any $n$-branch we have $\delt<T$:
Once $\delt$ crosses $T$, the free cell can ``squeeze in" an additional spike and the solutions bifurcate into a $(n+1)-(n+1)$ burst.

Distinguishing between the active and silent phases of a cycle allows us to describe the dynamics of the depression variable $d$ explicitly for each phase.
As can be seen from \cref{fig:free-quiet1}C, during the active phase $d$ depresses during the active phase of spikes and recovers during the inactive phases of spikes.
% TODO need to sort out terminology here, i.e. inactive vs quiet
In contrast, during the silent phase $d$ only recovers and does not depress.
Given the initial $d^{\star}=d(0)$ at the beginning of the cycle and the number of spikes in the free phase $n$, we can now construct the burst map $\Pi_{n}$.
The map
\begin{equation}
  \Pi_{n}(d^{\star})=Q_{n}\big(F_{n}(d^{\star}\big))
\end{equation}
\noindent
is a composition of two maps.
Map
\begin{equation}
  F_{n}:d^{\star}\mapsto \delt
\end{equation}
models the evolution of $d$ in the free phase.
$F_{n}$ takes an initial value $d^{\star}$ and calculates the inter-burst-interval $\delt$.
Map
\begin{equation}
  Q_{n}:\delt \mapsto d^{\star}
\end{equation}
models the recovery of $d$ in the quiet phase.
Given some $\delt$ map $Q_n$ computes $\dstar$ at the start of the next cycle.

@@

% TODO: Edit par
We start the construction of $\Pi_n$ by first considering the free phase and building the map $F_n$.
At each spike time $t_{k}$ where $d(t_k) = d_k$, variable $d$ decays first for the duration of the active phase of the spike for $T_{act}$, as described by the solution to \cref{eq:dot-d}.
At $t = t_k + T_{act}$ we have
\begin{equation}
  d(t_k + T_{act}) = d_k e^{-T_{act}/\tau_b}.
\end{equation}
The depression variable then recovers during the inactive phase of the spike until $t_{k+1}$, where for $0<t<T_{inact}$ we get
\begin{equation}
  d(t_{k+1}) = 1 - (1 - d_k e^{-T_{act}/\tau_b} )e^{-t/\tau_a}.
\end{equation}
By substituting $t=T_{inact}$ we can build a linear map that models the depression of $d$ from spike time $t_{k}$ to the subsequent spike time $t_{k+1}$ during the free phase:
\begin{equation}
  d_{k+1} = \lambda\rho d_{k} + (1-\rho),~\label{eq:map-d}
\end{equation}
where to keep the notation simple we let
\begin{align}
  \lambda & = \exp(-T_{act}/\tau_b),    \\
  \rho    & =\exp(-T_{inact}/\tau_{a}).
\end{align}

Given constant $T_{act}$ and $T_{inact}$, parameter $\lambda$ determines how much the synapses depresses during the active phase of the spike, while $\rho$ determines how much it recovers during the inactive phase.
Since $0<\lambda, \rho<1$, map \cref{eq:map-d} is increasing and contracting, with a fixed point at
\begin{equation}
  ~\label{eq:dsup}
  d_{s}=\frac{1-\rho}{1-\lambda\rho},
\end{equation}
where $0<d_{s}<1$.
The value $d_{s}$ is the maximum depression value that can be observed in the suppressed solution where the active cell fires at its uncoupled period $T$ (see \cref{fig:burst-sols}E).
Using the release condition in \cref{eq:release} allows us to derive the value of the the minimum coupling strength that will produce the full suppressed solution, denoted as $\gbar_{s}$.
% TODO Explain why s(0) = ds
% TODO See https://trello.com/c/AYqZIaXh: Would be good to elaborate on the idea that lambda and tau can be
% viewed as the new time constants in the reduced system.
Solving \cref{eq:dot-s} for $s(t)$ with $t=T$ and setting the initial value $s(0)=d_{s}$ gives
\begin{equation}
  \gbar_{s} d_{s}e^{-T/\tau_\kappa}=g^{\star}.
\end{equation}
By further substituting the definition of $d_{s}$ in~\eqref{eq:dsup} and rearranging, we can write $\gbar_{s}$ as a function of $\lambda$ and $\rho$:
\begin{equation}
  \label{eq:gs}
  \gbar_{s}(\lambda, \rho) = g^{\star}e^{T/\tau_\kappa}\frac{1-\lambda \rho}{1-\rho}.
\end{equation}
Note that the above dependence of $\gbar_{s}$ on $\lambda$ is linear and monotonically decreasing.
Increasing $\lambda$ reduces the strength of the depression of the free cell.
This in turn allows the free cell to fully suppress the quiet cell at smaller values of $\gbar$.

Solving~\eqref{eq:map-d} gives us the linear map $\delta_{n}:d^{\star}\mapsto
d_{n}$, that for some initial $d^{\star}$ computes the depression at the $n$th
spike time, $d_{n}=d(t_{n}^{-})$:
\begin{equation}
  ~\label{eq:delta-map}
  \delta_{n}(d^{\star}) = (\lambda \rho)^{n-1} d^{\star} + (1-\rho)\sum_{i=0}^{n-2}(\lambda \rho)^{i}.
\end{equation}
Since $\lambda < 1$, function $\delta_{n}$ is a linearly increasing function of
$d^{\star}$ with a fixed point at $d_{s}$ for all $n$.  Having identified $d$ after $n$
spikes, we can now use the release condition $\gbar s = g^{\star}$ (\cref{eq:release}) to
find $\delt$.
% FIXME:
After the last spike of the free phase at time $t_{n} = (n-1)T$ the synapse variable $s$
has the value of $d$ for the duration of $T_{act}$, which is given by
$\delta_n(\dstar)\lambda$. $s$ then decays exponentially for $\Delta t < T_{inact}$.
Solving \cref{eq:dot-s} (case $v < v_\theta$) with initial condition
$s(0)=\dn(\dstar)\lambda$ yields:
\begin{equation}
  ~\label{eq:s-sol}
  s(\delt)=\dn(\dstar)e^{-\Delta t/\tau_\kappa}.
\end{equation}
Substituting $s(\delt)$ into $s$ of the release condition (\cref{eq:release})
gives then
\begin{equation}
  ~\label{eq:release2}
  \gbar \dn(\dstar) e^{-\delt/\tau_\kappa}=g^{\star}.
\end{equation}
Our assumption of the release condition guarantees that the quiet cell 2 spikes and
becomes free when $\gbar s - g^{\star}$ crosses zero. Solving \cref{eq:release2} for
$\delt$ allows us to compute the inter-spike-interval as a function of $\dstar$, which
defines our map $F_n$:
\begin{equation}
  ~\label{eq:Fn-map}
  F_{n}(d^{\star}):=\tau_\kappa \ln{\left(\frac{\gbar }{g^{\star}} \delta_{n}(d^{\star})\right)}= \delt.
\end{equation}

@@

From \cref{eq:Pn-map} it is evident that $\Pi_{n}$ is monotonically increasing with respect to $\gbar$ and also $\dstar$:
\begin{align}
  \label{eq:non-degen1}
  \dv{\Pi_{n}}{\gbar}  & >0, \\[0.5ex]
  \label{eq:dPidd}
  \dv{\Pi_{n}}{\dstar} & >0,
\end{align}
and in the following sections we will heavily rely on this monotonicity property of $\Pi_n$.
Just as $F_{n}$, curves $\Pi_{n}$ spawn at the asymptote $d_{a}$ (\cref{eq:da}), and because
\begin{equation}
  \lim_{\gbar \to \infty}\Pi_{n} = 1\text{ for all }n,
\end{equation}
fixed points of $\Pi_{n}$ lie in $(d_{a}, 1)$.

% \subsection{Existence and stability of fixed points of $\mathit{\Pi_{n}}$}
\subsection{Existence and stability of fixed points}
We introduce the fixed point notation $\dstar_{f}$ with $\Pi_{n}(\dstar_{f})=\dstar_{f}$.
The existence of fixed points $\dstar_{f}$ for $\gbar$ sufficiently large can be shown from the strict monotonicity of $\Pi_{n}$ with respect to $\gbar$ and $\dstar$ (\cref{eq:dPidd,eq:non-degen1}), as well as the fact that the slope of $\Pi_{n}$ is monotonically decreasing,
\begin{equation}
  \label{eq:non-degen2}
  \left(\frac{\mathrm{d}}{\mathrm{d}\dstar}\right)^2 \Pi_{n}<0.
\end{equation}
In the limit $\dstar \to d_{a}$ the value of $\Pi_n$ decreases without bound for any $\gbar>0$.
In the limit $\gbar \to 0$, $\Pi_n$ also decreases without bound, but as $\gbar \to \infty$ values of $\Pi_n$ approach $1$.
It follows from \cref{eq:non-degen1} and the intermediate value theorem that for some $\gbar$ large enough $\Pi_n$ intersects the diagonal.
Moreover, because $\Pi_n$ and its slope are monotonic with respect to $\dstar$, there exists some critical fixed point $(\dstar_b, \gbar_b)$ where $\Pi_n$ aligns with the diagonal tangentially with
\begin{align}
  \Pi_{n}(\dstar_{b}; \gbar_{b})                     & =\dstar_{b}, \\
  \frac{\d}{\d \dstar}\Pi_{n}(\dstar_{b}; \gbar_{b}) & =1.
\end{align}
Equations~\eqref{eq:non-degen1} and~\eqref{eq:non-degen2} constitute the non-degeneracy conditions for a codimension-1 fold bifurcation of maps, indicating that in a neighbourhood of $(\dstar_{b}, \gbar_{b})$ map $\Pi_{n}$ has the topological normal form described by the graph of
\begin{equation}
  \label{eq:normal-form}
  x\mapsto \beta+x-x^{2},
\end{equation}
with a stable and unstable fixed point $x=\pm\sqrt{\beta}$, and slopes $\dv*{x}{\beta}=\mp {(2\sqrt{\beta})}^{-1}$, respectively.
